Global Distribution of Unintended Pregnancies and Contraceptive Usage

### 2024-03-14 code/analysis-code/statistical-analysis.qmd
Manuscript
Finished describing how to reproduce the project up to this point and added the statistical analysis in the manuscript.

# save fit results table  
table_file1 = here("results", "resulttable1.rds")
saveRDS(lmtable1, file = table_file1)
```
There appears to be a very strong relationship between total number of women using this birth control method and the outcome based on the coefficient estimate of 87.65998.

@@

# save fit results table  
table_file2 = here("results", "resulttable2.rds")
saveRDS(lmtable2, file = table_file2)
```
This model shows that the Latin America/the Carribean region has the strongest association with the outcome variable based on a coefficient estimate of 1.6218844 followed closely by the Asia region. The Oceania region has the weakest association with the outcome variable based on a coefficient estimate of 0.2512100.

@@

# save fit results table  
table_file3 = here("results", "resulttable3.rds")
saveRDS(lmtable3, file = table_file3)
```
This model shows there is a negative relationship between the number of women currently married and the outcome. As the percentage of currently married women increases by 1, the percentage of unintended pregnancies decreases by a factor of 0.4181571.

@@

# save fit results table  
table_file4 = here("results", "resulttable4.rds")
saveRDS(lmtable4, file = table_file4)
```
This model shows a positive relationship between the percentage of women never married and the outcome. As the percentage of women never married increases by 1, the percentage of unintended pregnancies increases by a factor of 0.2474218.

@@

# save fit results table  
table_file5 = here("results", "resulttable5.rds")
saveRDS(lmtable5, file = table_file5)
```
This model shows that total rate of maternal deaths is slightly negatively correlated with the outcome based on a small coefficient estimate of -0.01482568.

@@

# save fit results table  
table_file6 = here("results", "resulttable6.rds")
saveRDS(lmtable6, file = table_file6)
```
This model shows that total rate of maternal deaths due to unintended pregnancies has an even weaker association with the outcome based on a small coefficient estimate of -0.003158542.

@@

# save fit results table  
table_file7 = here("results", "resulttable7.rds")
saveRDS(lmtable7, file = table_file7)
```
This model shows that total rate of maternal deaths due to abortion has the strongest association with the outcome out of the three maternal mortality rates included based on a coefficient estimate of -0.033353527.

@@

# save fit results table  
table_file8 = here("results", "resulttable8.rds")
saveRDS(lmtable8, file = table_file8)
```
This model shows the strongest predictor is the number of women using short acting reversible contraception with a huge coefficient of 85.6433. The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor. The Oceania region now has a negative association with the outcome. Both of the marriage status predictors are associated with a decrease in the outcome. The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both flipped from the single predictor models to show a positive association with the outcome.

>>>>

# save fit results table  
table_file1 = here("results", "tables", "resulttable1.rds")
saveRDS(lmtable1, file = table_file1)
```
There appears to be a very strong relationship between total number of women using this birth control method and the outcome based on the coefficient estimate of 87.65998.

@@

# save fit results table  
table_file2 = here("results", "tables", "resulttable2.rds")
saveRDS(lmtable2, file = table_file2)
```
This model shows that the Latin America/the Carribean region has the strongest association with the outcome variable based on a coefficient estimate of 1.6218844 followed closely by the Asia region. The Oceania region has the weakest association with the outcome variable based on a coefficient estimate of 0.2512100.

@@

# save fit results table  
table_file3 = here("results", "tables", "resulttable3.rds")
saveRDS(lmtable3, file = table_file3)
```
This model shows there is a negative relationship between the number of women currently married and the outcome. As the percentage of currently married women increases by 1, the percentage of unintended pregnancies decreases by a factor of 0.4181571.

@@

# save fit results table  
table_file4 = here("results", "tables", "resulttable4.rds")
saveRDS(lmtable4, file = table_file4)
```
This model shows a positive relationship between the percentage of women never married and the outcome. As the percentage of women never married increases by 1, the percentage of unintended pregnancies increases by a factor of 0.2474218.

@@

# save fit results table  
table_file5 = here("results", "tables", "resulttable5.rds")
saveRDS(lmtable5, file = table_file5)
```
This model shows that total rate of maternal deaths is slightly negatively correlated with the outcome based on a small coefficient estimate of -0.01482568.

@@

# save fit results table  
table_file6 = here("results", "tables", "resulttable6.rds")
saveRDS(lmtable6, file = table_file6)
```
This model shows that total rate of maternal deaths due to unintended pregnancies has an even weaker association with the outcome based on a small coefficient estimate of -0.003158542.

@@

# save fit results table  
table_file7 = here("results", "tables", "resulttable7.rds")
saveRDS(lmtable7, file = table_file7)
```
This model shows that total rate of maternal deaths due to abortion has the strongest association with the outcome out of the three maternal mortality rates included based on a coefficient estimate of -0.033353527.

@@

# save fit results table  
table_file8 = here("results", "tables", "resulttable8.rds")
saveRDS(lmtable8, file = table_file8)
```
This model shows the strongest predictor is the number of women using short acting reversible contraception with a huge coefficient of 85.6433. The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor. The Oceania region now has a negative association with the outcome. Both of the marriage status predictors are associated with a decrease in the outcome. The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both flipped from the single predictor models to show a positive association with the outcome.

### 2024-03-14 products/manuscript/Manuscript.qmd
Manuscript
Finished describing how to reproduce the project up to this point and added the statistical analysis in the manuscript.

# Research Questions
1) Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among countries? 
    The main outcome variables will be total percentage of unintended pregnancies among women using short acting          reversible contraception methods (`pct_upreg_sarc`).

@@

1. Is there a significant difference in the percentage of unplanned pregnancies between three contraceptive methods and no usage between regions of the world?

@@

The next step is exploratory data analysis. The eda.qmd file in the code > eda-code folder contains all the code for this process. We produce a couple summary tables before we begin a simple exploration of trends in percentage of unintended pregnancies by country in our first set of histograms. The second set of histograms explore contraceptive usage methods by country. These graphs revealed outliers, so we standardized by population size for each country. The next sets of plots are line graphs showing total usage of contraceptive methods to percentage of unintended pregnancies associated with that method. All of these figures are saved in the Supplementary-Material.qmd file in the products > manuscript > supplement folder. The last two sets of plots contain line graphs of the relationship between the two variables stratified by region and violin plots displaying the relationship between percentage of unintended pregnancies and the categorical usage variable created during processing, which are included in this manuscript. The exploratory data analysis process led to another version of the processed data with standardized total usage variables saved as processeddata2.rds in the data > procsssed-data folder.

Step three is the statistical analayis and model fitting portion found in the statistical-analysis.qmd file in the code > analysis-code folder.

@@

## Statistical analysis
_Explain anything related to your statistical analyses._

@@

shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
# knitr::include_graphics(here("results","height-weight-stratified.png")) #commented out for the sake of knitting our project at this stage
```

>>>>

# Research Questions
1) Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among countries?

The main outcome variables will be total percentage of unintended pregnancies among women using short acting          reversible contraception methods (`pct_upreg_sarc`).

@@

1) Is there a significant difference in the percentage of unintended pregnancies among short acting reversible contraception users among countries?

The main outcome variables will be total percentage of unintended pregnancies among women using short acting          reversible contraception methods (`pct_upreg_sarc`).

@@

The next step is exploratory data analysis. The eda.qmd file in the code > eda-code folder contains all the code for this process. We produced a couple summary tables before we began a simple exploration of trends in percentage of unintended pregnancies by country in our first set of histograms. The second set of histograms explore contraceptive usage methods by country. These graphs revealed outliers, so we standardized by population size for each country. The next sets of plots are line graphs showing total usage of contraceptive methods to percentage of unintended pregnancies associated with that method. The last two sets of plots contain line graphs of the relationship between the two variables stratified by region and violin plots displaying the relationship between percentage of unintended pregnancies and the categorical usage variable created during processing. All of these figures are saved in the Supplementary-Material.qmd file in the products > manuscript > supplement folder. This process led us to focus on the percentage of unintended pregnancies among short acting reversible methods as our single outcome because this method is becoming increasingly popular, so we feel that results from this contraception method will be the most relevant.

Step three is the statistical analysis and model fitting portion found in the statistical-analysis.qmd file in the code > analysis-code folder. We explore the single outcome `pct_upreg_sarc` based on 7 predictors with simple linear regression, total number of women using short acting reversible methods, region, percent of currently married women, percent of women who never married, total rate of maternal mortality, rate of maternal mortality by unintended pregnancy, and rate of maternal mortality by abortion. We also assess the performance of the models in this file. Result tables for each model are stored in the results > tables folder.

@@

## Statistical analysis
_Explain anything related to your statistical analyses._
We began with a simple linear regression model with each of the 7 predictor variables. We also included a multiple linear regression model with all 7 predictors. We included the RMSE for two of the models to begin exploring model fit.

@@

We completed a simple linear regression analysis with each of the 7 predictors. The predictor with the strongest relationship with the outcome is total number of women using short acting reversible methods (`sarc_standard`).

```{r}
#| label: basic-statistical-analysis
#| fig-cap: "Strongest Predictor: sarc_standard"
#| echo: FALSE
readRDS(here("results","tables", "resulttable1.rds")) 
```
The model using the region predictor revealed that the Latin America/the Caribbean region had the strongest relationship with the outcome variable followed closely by the Asia region. The Oceania region does not appear to have a strong effect on the outcome variable with the Europe region only showing a slightly larger association.
```{r}
#| label: basic-statistical-analysis
#| fig-cap: "Regional Predictor"
#| echo: FALSE
readRDS(here("results","tables", "resulttable2.rds")) 
```
We created one model with all 7 predictors using multiple linear regression. This model shows the strongest predictor is the number of women using short acting reversible contraception. The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor. The Oceania region now has a negative association with the outcome. Both of the marriage status predictors are associated with a decrease in the outcome. The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both flipped from the single predictor models to show a positive association with the outcome. 
```{r}
#| label: basic-statistical-analysis
#| fig-cap: "Multiple Linear Regression with 7 Predictors"
#| echo: FALSE
readRDS(here("results","tables", "resulttable8.rds")) 
```
The full statistical analysis can be found in the statistical-analysis.qmd file in the code > analysis-code folder.
